The Splashback Radius of Planck-SZ clusters

### 2019-02-14 Paper_v2/Reply_to_referee.txt

Reply: Our original motivation to keep Figure 1 in the beginning was to show the
average redshift of the sample and the mass range, since that goes in our
theoretical expectation. However since this is indicated in the Table, we have
now moved Figure 1 to the appendix based on the referee's suggestion.

>
> b) Figures 2 and 3 show very similar things, namely the 2D and 3D density
> profiles. However, because the figures are split and because the panels come
> out much larger in Figure 3, it is not immediately apparent that the actual
> measured results are shown in Figure 2. I have a somewhat radical suggestion: I
> would combine both figures by stacking the panels vertically instead of
> horizontally, i.e., combining the 2D and 3D profiles into one column of five
> panels that shows all aspects of the modeling (data, inferred 2D, 2D slope, 3D
> etc). Then, the three samples would be three columns that could easily be fit
> next to each other, particularly as the y-labels would not need to be repeated
> in each column. The disadvantage is that it would become harder to compare the
> location of Rsp in the three samples, but as stated in the text, they all agree
> anyway. The advantage would be that one could directly compare the inferred 2D
> and 3D radii -- to me, that is more informative. Most importantly, it would be
> very clear that the combined figure represents the main result of the paper.
> The same could be done for Figures 6 and 7 which, again, appear somewhat
> disconnected to the reader even though they are highly connected.

Reply: We thank the referee for this suggestion. We have now combined Figure 2
and Figure 3 in the original manuscript into one. However, since the upper two
panels have the projected radius as the abscissa while the bottom three have
the 3-dimensional radius instead, we have included the labels at the bottom of
row 2, and row 5 separately. We have also included a label Rsp^{\rm 2d} and
Rsp^{\rm 3d} to indicate the two and three dimensional locations of the
steepest slopes, since it could be confusing. We have similarly combined Figure
6 and 7 into the new Figure 4.

@@

>
> b) In Section 4, you explain that the splashback radius seen in the blue
> population may be due to contamination of the blue sample, but the conclusion
> remains a little unclear in the main text. I would refer to Baxter+17 Figure 5,
> saying that they found no splashback in the blue population, and make a firm
> statement on whether there is any evidence that this is wrong (given the issues
> with contamination).  -refer to baxter and explain why we find something
> different
>
> 4) I have some smaller comments and questions regarding the text and figures.
>
> a) In Table 2, it looks like "'" were used instead of "." in large numbers.

Reply: We have corrected this.

>
> b) The derivatives in figures 2, 3 etc look somewhat noisy, even though the
> profile measure is smooth as it is inferred from the profile fitting function.
> My expectation would be that the derivatives should look smooth too?

Reply: The noisy looking derivatives were a result of us avoiding the large
computational cost of storing the derivatives at a large number of radial
points at every MCMC step. We have now taken the computed points, and plotted
the spline instead and are displaying it in the Figure to avoid this noise in
the derivative figure.

### 2019-02-14 Paper_v2/RSP_paper_new.tex
more edits

\begin{abstract}
We present evidence for the existence of the splashback radius in galaxy
clusters selected using the Sunyaev-Zeldovich effect, a sample unaffected by
systematics related to cluster finding in the optical wavelength range. We show
that the deprojected cross-correlation of galaxy clusters found in the Planck
survey with galaxies detected photometrically in the Pan-STARRS survey, shows a
sharp steepening feature (a logarithmic slope steeper than $-3$), which we
associate with the splashback radius. We infer the three-dimensional splashback
radius for the SZ cluster sample to be $r_{\rm sp}=1.85_{-0.30}^{+0.26}$
$\mpch$, where the cluster sample has an average halo mass $M_{\rm
500c}=3.0\times10^{14}\msunh$ at an average redshift of $z=0.18$.  The inferred
value of the splashback radius appears consistent with the expected location
for dark matter halos in the standard cold dark matter paradigm. However, given
the limited precision of our measurements, we cannot conclusively confirm or
rule out the smaller splashback radius measured so far in the literature for
optically selected galaxy clusters. We show that the splashback radius does not
depend upon the galaxy magnitude for galaxies fainter than $M_i-5\log
h=-19.44$, and is present at a consistent location in galaxy populations
divided by color. The presence of the splashback radius in the star-forming
galaxy population could potentially be used to put lower limits on the
quenching timescales for galaxies. We can marginally rule out the contamination
of the star-forming galaxy sample by quenched galaxies, but the results would
need further verification with deeper datasets.

@@

\textbf{Dynamical friction is expected to slow down massive subhalos, hosting
galaxies that are orbiting the galaxy cluster \citep{chandrasekhar1943dynamical}.
If a majority of the observed galaxies are hosted by these massive
subhalos, the slow down due to dynamical friction would reduce the apocenters
of their orbits, and in turn, biasing the measured splashback radius of the host
cluster \citep{more2016detection, diemer2017sparta}. As dynamical
friction timescales are inversely proportional to the mass of the subhalos,
this effect is stronger for more massive subhalos. Therefore, one expects the
splashback radius to decrease for more massive subhalos
\citep{adhikari2016observing}.  This bias can be limited by choosing a faint
galaxy sample such that the majority of the galaxies reside in low mass subhalos,
unaffected by dynamical friction.} The SDSS sample used by
\citet{more2016detection} was already deep enough such that biases in the
location of the splashback radius due to dynamical friction effects were
expected to be small. Nevertheless, we use galaxy samples, which are fainter by
$0.5-1$ magnitudes compared to those used by \citet{more2016detection}.

@@

Therefore the model is described by eight model parameters.
We infer the posterior distributions of those parameters by fitting the model for the two-dimensional correlation
function to the measured two-dimensional correlation signal.
We use the affine invariant Markov Chain Monte Carlo sampler of \citet{goodman2010ensemble} as
implemented in the parallel python package {\it emcee} by
\citet{foreman2013emcee}.
\textbf{We let the parameters $\rho_{\mathrm{s}}$, $\rho_0$ and
$s_{\mathrm{e}}$ vary freely during our fitting procedure. Similar to
\citet{more2016detection} we put flat priors on $\log_{10}(r_{\mathrm{s}})$ and
$\log_{10}(r_{\mathrm{t}})$ but double their ranges to $[0.1,5.0]$ compared to
their work. In contrast to \citet{more2016detection}, we use a normal prior
with a central value of $\log_{10}(0.2)$ and a scale of 1.2 for
$\log_{10}(\alpha)$ as deduced from mass estimates \citep{gao2008redshift}.
For the parameters $\log_{10}(\beta)$ and $\log_{10}(\gamma)$ we also use
normal priors with central values $\log_{10}(6.0)$ and $\log_{10}(4.0)$,
respectively and scale 0.4 as recommended by \citet{diemer2014dependence}.}

@@

\textbf{The location of the splashback radius can be identified in various
ways, e.g., as the radius of the outermost shell in theoretical models of
spherical symmetric collapse \citep{adhikari2014splashback, shi2016outer}, or
using non-spherical shells of density jumps \citep{ mansfield2017splashback},
or the average location of the apocenter of particles falling in the cluster
potential \citet{diemer2017splashback}. Although these definitions have subtle
differences, they all follow the general trend of a decreasing splashback
radius with increasing rate of accretion \citep{diemer2017splashback}.
In this paper, we adopt the convention of
\citet{more2015splashback}, who define the location of the splashback radius
for halos on galaxy cluster scales to be the location of the steepest
logarithmic slope of the radially averaged density profile, due to its ease of
accessibility in data.}

@@

\section{Results}
\label{sec:Results}
The measurements of the two-dimensional cluster-galaxy
cross-correlations are shown as black points with error bars in the
first row of Figure~\ref{fig:all_graphs}. The different columns
correspond to the three different absolute magnitude limits that we
have used to select all the galaxies when calculating the
cross-correlations. The cross-correlation signal is clearly detected
in all three measurements. \textbf{The signal to noise ratios of the
three different estimates of the two-dimensional cross-correlation signals
amount to 42.4, 43.3 and 30.9 for the galaxy samples PS 21, PS 21.5 and PS 22,
respectively. For comparison, the detection SNR in \citet{more2016detection}
was 263, a much higher number due to the size of the cluster sample used in
their study.}

@@

\textbf{We show the two-dimensional posterior distributions for each pair of
parameters for the PS 21.5 sample in Figure~\ref{fig:corner_21.5}. The
posterior distributions for the other datasets show similar degeneracies and
are not included in the paper for brevity.
It is observed that there are some strong degeneracies present between
parameters describing the same part of the model in Equation~\ref{eq:model}
(namely $\rho_{\mathrm{in}}$, $f_{\mathrm{trans}}$ and $\rho_{\mathrm{out}}$).
This is expected since certain combinations of parameter values can
very well result in the same functional shape of the different parts of
the model. This degeneracy also leads to some of the parameters not being
constrained tightly (especially $\rho_s$ and $r_s$ due to their strong degeneracy
with $\alpha$). On the other hand, it is important to note that there is
little-to-no degeneracy present between parameters describing different parts
of the model, which means that the overall shape of the profile is well
constrained.  Note also that, since the location of the splashback radius only
depends on the overall shape of the profile, the uncertainties of the
individual parameters within the independent parts of the profile do not
translate into a huge uncertainty on the location of the splashback feature.}

\begin{figure*}
	\centering{
    \includegraphics[scale=0.85]{MAIN_results_with_spline_no_mc.pdf}}
\caption{The main results of the study are presented in this Figure.
The first two rows show the inferred two-dimensional cross-correlation
signals as well as its derivatives.
The three different columns correspond to the three different galaxy
samples cross-correlated with the cluster sample.
The magnitude limit applied to the galaxy catalog is indicated in
each tile of the first row along with the signal-to-noise ratio (SNR).
The measurements of the two-dimensional cross-correlation
signals (black dots) are shown in the first row as well. There, the colored
curves show the two-dimensional model fits of the functional
form in Equation~\ref{eq:surface}. The vertical, shaded regions indicate
the estimates of the locations of steepest slope of the profiles as estimated from the corresponding minima of the two-dimensional
derivative profiles, which are shown in the second row.
Rows three to five show the corresponding three-dimensional model fit
with the functional form Equation~\ref{eq:model} as well as its first and
second order derivatives.
In rows three to five the vertical, shaded regions indicate the
estimates of the splashback radii as estimated from the corresponding minima
of the three-dimensional derivative profiles.
The grey curves show the corresponding, estimated first and second derivatives of the
one-halo term (namely $\rho_{\mathrm{in}}(r)f_{\mathrm{trans}}(r)$ in Equation~\ref{eq:model}).
For comparison, the location of the $r_{\mathrm{200m}}$ radius as calculated
from the average cluster sample properties is indicated by the black, dotted
lines.}
   \label{fig:all_graphs}
\end{figure*}

>>>>

\begin{abstract}
We present evidence for the existence of the splashback radius in galaxy
clusters selected using the Sunyaev-Zeldovich effect, {\bf a sample unaffected by
systematics related to cluster finding in the optical wavelength range.} We show
that the deprojected cross-correlation of galaxy clusters found in the Planck
survey with galaxies detected photometrically in the Pan-STARRS survey, shows a
sharp steepening feature (a logarithmic slope steeper than $-3$), which we
associate with the splashback radius. We infer the three-dimensional splashback
radius for the SZ cluster sample to be $r_{\rm sp}=1.85_{-0.30}^{+0.26}$
$\mpch$, where the cluster sample has an average halo mass $M_{\rm
500c}=3.0\times10^{14}\msunh$ at an average redshift of $z=0.18$.  The inferred
value of the splashback radius appears consistent with the expected location
for dark matter halos in the standard cold dark matter paradigm. However, given
the limited precision of our measurements, we cannot conclusively confirm or
rule out the smaller splashback radius measured so far in the literature for
optically selected galaxy clusters. We show that the splashback radius does not
depend upon the galaxy magnitude for galaxies fainter than $M_i-5\log
h=-19.44$, and is present at a consistent location in galaxy populations
divided by color. The presence of the splashback radius in the star-forming
galaxy population could potentially be used to put lower limits on the
quenching timescales for galaxies. We can marginally rule out the contamination
of the star-forming galaxy sample by quenched galaxies, but the results would
need further verification with deeper datasets.

@@

\textbf{Dynamical friction is expected to slow down massive subhalos, hosting
galaxies that are orbiting the galaxy cluster \citep{chandrasekhar1943dynamical}.
If a majority of the observed galaxies are hosted by these massive
subhalos, the slow down due to dynamical friction would reduce the apocenters
of their orbits, and in turn, bias the measured splashback radius of the host
cluster \citep{more2016detection, diemer2017sparta}. As dynamical
friction timescales are inversely proportional to the mass of the subhalos,
this effect is stronger for more massive subhalos. Therefore, one expects the
splashback radius to be biased low as measured for more massive subhalos
\citep{adhikari2016observing}.  This bias can be overcome by choosing a faint
enough galaxy sample such that the majority of the galaxies reside in low mass
subhalos, unaffected by dynamical friction.} The SDSS sample used by
\citet{more2016detection} was already deep enough such that biases in the
location of the splashback radius due to dynamical friction effects were
expected to be small. Nevertheless, we use galaxy samples, which are fainter by
$0.5-1$ magnitudes compared to those used by \citet{more2016detection}.

@@

Therefore the model is described by eight model parameters.
We infer the posterior distributions of those parameters by fitting the model for the two-dimensional correlation
function to the measured two-dimensional correlation signal.
We use the affine invariant Markov Chain Monte Carlo sampler of \citet{goodman2010ensemble} as
implemented in the parallel python package {\it emcee} by
\citet{foreman2013emcee}.
\textbf{We let the parameters $\rho_{\mathrm{s}}$, $\rho_0$ and
$s_{\mathrm{e}}$ vary freely during our fitting procedure. Similar to
\citet{more2016detection} we put flat priors on $\log_{10}(r_{\mathrm{s}})$ and
$\log_{10}(r_{\mathrm{t}})$ but double their ranges to $[0.1,5.0]$ compared to
their work in order to stay conservative. We also use a normal prior with a
central value of $\log_{10}(0.2)$ for $\log_{10}(\alpha)$ as deduced from mass
estimates \citep{gao2008redshift}. To be conservative, we once again double the
scale of the prior to be 1.2 compared to \citet{more2016detection}. For the
parameters $\log_{10}(\beta)$ and $\log_{10}(\gamma)$ we also use normal priors
with central values $\log_{10}(6.0)$ and $\log_{10}(4.0)$, respectively and
scale 0.4 as recommended by \citet{diemer2014dependence}.}

@@

\textbf{The location of the splashback radius can be identified in various
ways, e.g., as the radius of the outermost shell in theoretical models of
spherical symmetric collapse \citep{adhikari2014splashback, shi2016outer}, or
using non-spherical shells of density jumps \citep{ mansfield2017splashback},
or the average location of the apocenter of particles falling into the cluster
potential \citet{diemer2017splashback}. Although these definitions have subtle
differences, they all follow the general trend of a decreasing splashback
radius with increasing rate of accretion \citep{diemer2017splashback}.
In this paper, we adopt the convention of
\citet{more2015splashback}, who define the location of the splashback radius
for halos on galaxy cluster scales to be the location of the steepest
logarithmic slope of the radially averaged density profile, due to its ease of
accessibility in data.}

@@

\section{Results}
\label{sec:Results}
The measurements of the two-dimensional cluster-galaxy
cross-correlations are shown as black points with error bars in the
first row of Figure~\ref{fig:all_graphs}. The different columns
correspond to the three different absolute magnitude limits that we
have used to select all the galaxies when calculating the
cross-correlations. The cross-correlation signal is clearly detected
in all three measurements. \textbf{The signal-to-noise ratios of the
three different estimates of the two-dimensional cross-correlation signals
amount to 42.4, 43.3 and 30.9 for the galaxy samples PS 21, PS 21.5 and PS 22,
respectively. For comparison, the detection signal-to-noise ratio in
\citet{more2016detection} was 263, a much higher number due to the size of the
cluster sample used in their study.}

@@

\textbf{We show the two-dimensional posterior distributions for each pair of
parameters for the PS 21.5 sample in Figure~\ref{fig:corner_21.5}. The
posterior distributions for the other datasets show similar degeneracies and
are not included in the paper for brevity.
It is observed that there are some strong degeneracies present between
parameters describing the same part of the model in Equation~\ref{eq:model}
(namely $\rho_{\mathrm{in}}$, $f_{\mathrm{trans}}$ and $\rho_{\mathrm{out}}$).
This is expected since certain combinations of parameter values can
very well result in the same functional shape of the different parts of
the model. This degeneracy also leads to some of the parameters not being
constrained tightly (especially $\rho_s$ and $r_s$ due to their strong degeneracy
with $\alpha$). On the other hand, it is important to note that there is
little-to-no degeneracy present between parameters describing different parts
of the model, which means that the location of the transition region is well
constrained. Therefore, the uncertainties of the individual parameters within
the independent parts of the profile do not directly translate into an
uncertainty in the location of the splashback feature.}

\begin{figure*}
	\centering{
    \includegraphics[scale=0.75]{MAIN_results_with_spline_no_mc.pdf}}
\caption{The main results of the study are presented in this Figure.
The first two rows show the inferred two-dimensional cross-correlation
signals as well as its derivatives.
The three different columns correspond to the three different galaxy
samples cross-correlated with the cluster sample.
The magnitude limit applied to the galaxy catalog is indicated in
each tile of the first row along with the signal-to-noise ratio (SNR).
The measurements of the two-dimensional cross-correlation
signals (black dots) are shown in the first row as well. There, the colored
curves show the two-dimensional model fits of the functional
form in Equation~\ref{eq:surface}. The vertical, shaded regions indicate
the estimates of the locations of steepest slope of the profiles as estimated from the corresponding minima of the two-dimensional
derivative profiles, which are shown in the second row.
Rows three to five show the corresponding three-dimensional model fit
with the functional form Equation~\ref{eq:model} as well as its first and
second order derivatives.
In rows three to five the vertical, shaded regions indicate the
estimates of the splashback radii as estimated from the corresponding minima
of the three-dimensional derivative profiles.
The grey curves show the corresponding, estimated first and second derivatives of the
one-halo term (namely $\rho_{\mathrm{in}}(r)f_{\mathrm{trans}}(r)$ in Equation~\ref{eq:model}).
For comparison, the location of the $r_{\mathrm{200m}}$ radius as calculated
from the average cluster sample properties is indicated by the black, dotted
lines.}
   \label{fig:all_graphs}
\end{figure*}
